Circumpolar variations in Southern Ocean eddy dynamics:\\An ensemble approach

### 2021-05-26 paper/main.tex
more paper updates
editing

\section{Results}
Figure 1 - (a) TKE snapshot and (b) TKE climatology and perhaps (c) ensemble spread in TKE.

Figure 2 - Entire Southern Ocean:(a,b) cherrypicked experiments TKE + lag; (c) ensemble mean TKE + lag

Figure 3 - 4 or 6 sectors TKE + lag

Figure 4 - interesting regions - again 4 or 6 of them, TKE + lags

Figure 5 - Ri ratio plot, and other global features

>>>>

\section{Results}

The intensity of the Southern Ocean eddy field is not uniform.
Snapshots of eddy kinetic energy (e.g. Fig.~\ref{Fig:1}a) show the occurrence of strong eddies which occur in the lee of subsurface topography and at the outlet from western boundary currents such as the Agulhas retroflection and Malvinas current.
The same patterns are evident in the ensemble mean of EKE (Fig.~\ref{Fig:1}b), although the signal of individual eddies is not longer apparent.
The strongest band of EKE approximately follows the path of the Antarctic Circumpolar Current, and EKE is generally weak south of 60$^\circ$S.
The patterns of EKE in this model broadly match the regional variations of EKE observed from satellite altimetry (REF), albeit at slightly lower intensity, as expected in an eddy-permitting model (e.g. Kiss et al. REF).

\begin{figure}[ht]
\begin{center}
\includegraphics[width=\hsize]{Figure1}
\caption{(a) EKE snapshot and (b) ensemble mean EKE  }
\label{Fig:1}
\end{center}
\end{figure}

For each of the 50 members of the OCCIPUT ensemble we take the EKE averaged over the entire Southern Ocean, and plot the timeseries of this quantity in thin grey lines in Fig.~\ref{Fig:2}(a). 
This plot highlights the considerable spread in EKE, even when averaged over the full circumpolar belt; in other words, there is a significant component of intrinsic variability in the Southern Ocean eddy field.
Nonetheless, when averaged over all ensemble members (red line in Fig.~\ref{Fig:2}a) the coherent, forced, component of eddy variability is revealed.
Averaged over this region the fraction, $R_i$, of intrinsic variance is 0.69, confirming the visual impression that intrinsic processes are dominant here.

\begin{figure}[t]
\begin{center}
\includegraphics[width=\hsize]{Figure2}
\caption{Eddy kinetic energy statistics over the Southern Ocean (40$^\circ$S--60$^\circ$S). (a) Spatially averaged EKE for individual ensemble members (grey) and the ensemble mean (red) along with wind stress (black); and (b) Time-lagged correlation for the ensemble mean (blue) and individual ensemble  members (grey) -- with two individual ensemble members highlighted in magenta/orange.}
\label{Fig:2}
\end{center}
\end{figure}

Although Southern Ocean EKE is strongly intrinsic, there remains a significant component of forced variability.
Previous studies have suggested that there is a strong contribution of wind stress forcing upon EKE, and we therefore compare the forced variability with the variations in wind stress averaged over the same circumpolar belt (black line in Fig.~\ref{Fig:2}a).
This comparison suggests a relationship in which wind stress leads variations in EKE, consistent with previously published results.
However, the time-lagged correlations between wind stress and EKE suggests that this relationship is complex. 
The intrinsically variable nature of Southern Ocean eddies means that for some ensemble members, there is no meaningful correlation between wind stress and eddies (grey lines in Fig~\ref{Fig:2}b).
On the other hand, ensemble member 4 (magenta line) has a clear an significant (??) correlation with a 7-month lag, while member 5 (orange line) is correlated weakly at 6 months, and strongly correlated at a 28-month lag.
These isolated examples highlight how each ensemble member behaves differently, but the ensemble mean (blue line) includes a correlation at $\sim$6 months and a second clear peak at $\sim$30 months.
Thus, there appear to be two distinct timescales of response of the Southern Ocean eddy field to wind stress.

\begin{figure}[t]
\begin{center}
\includegraphics[width=\hsize]{Figure3}
\caption{Eddy kinetic energy statistics within sub-regions of the Southern Ocean. (a) Map showing Ensemble Mean EKE, along with 3 boxes over which a regional EKE analysis is applied; (b) Regional analysis of the Southeast Indian Ocean showing ensemble mean EKE in cyan, individual ensemble members EKE in grey and local wind stress averaged over the region in black; (c) Lagged correlation of ensemble mean EKE in each of the three regions; (d) Regional analysis of the Southwest Pacific Ocean showing ensemble mean EKE in red, individual ensemble members EKE in grey and local wind stress averaged over the region in black; and (3) Regional analysis of the South Atlantic Ocean showing ensemble mean EKE in orange, individual ensemble members EKE in grey and local wind stress averaged over the region in black. The fraction of intrinsic variance in each region is shown in the caption of panels (b), (d) and (e).}
\label{Fig:3}
\end{center}
\end{figure}

The ensemble of simulations shown here allow us to look in more detail at smaller regions of the Southern Ocean.
Calculating the variability of EKE in a smaller region has the advantage of isolating different processes which may occur in differing regions (for example, stronger topographic steering in places with steep bathymetry).
However, this advantage is partly offset by higher intrinsic variability in cases where the region of interest is so small that an individual eddy or event can have a large influence over the EKE timeseries.
In balancing these competing issues, we look at the variability within regions that span 15-20$\circ$ in latitude and 30-40$^\circ$ in longitude, as shown in Figure \ref{Fig:3}(a).
We analyse the EKE timeseries averaged over these boxes -- including individual member EKE, ensemble mean EKE and lagged correlations between local wind stress forcing and the ensemble mean EKE in Fig. \ref{Fig:3}(b-e).

We first examine a region in the lee of Kerguelen Plateau in the Southeast Indian Ocean (cyan box in Fig.~\ref{Fig:3}a).
This region is characterised by high-frequency ($\sim$1 year) variations in EKE, with a relatively large forced component (the intrinsic variance fraction $R_i = 0.51$ which is smaller than the Southern Ocean average; Fig.~\ref{Fig:3}b).
The forced variation is clearly evident in the timeseries of from individual ensemble members; and this forced component is closely related to wind stress.
The lag between wind stress variations and ensemble mean EKE is short ($\sim$ 6 months; Fig.~\ref{Fig:3}c) with a single peak in the lagged correlation.
This region highlights a regime in which the eddy field responds rapidly to variations in the local wind stress.

In the Southeast Pacific Ocean (red box in Fig.~\ref{Fig:3}a) the situation is clearly different.
Here, the intrinsic variance fraction is similar to the Southeast Indian Ocean ($R_i=0.56$) but the timescale of the variability is much longer and there is no significant response to wind stress variations which occur at sub-annual scales (Fig.~\ref{Fig:3}d).
The peak in the lag correlation occurs at approximately 14 months in this region, and again has only a single clear peak.
Thus, this region varies slowly and consistently to multi-year variations in wind stress.

in the South Atlantic Ocean (orange box in Fig.~\ref{Fig:3}a) the system is more dominated by intrinsic variance ($R_i=0.65$) and is poorly correlated with wind stress forcing (Fig.~\ref{Fig:3c,e), reinforcing the circumpolar variability of the EKE response to wind.

Figure 5 -- showing that it is local wind forcing that matters

Figure 6 - Ri ratio plot, and other global features
